---
title: "Dataframe Exercies"
subtitle: "Data 180, Professor Kessler"
author: 
  name: "My Name"
  email: "xxxx@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document contains a set of exercises which we will complete both in class and at home. Please attempt each of these exercises at your own pace.

### First steps...

Create a data frame named `students` with the following columns:
- `Name`: "Alice", "Bob", "Charlie", "David", "Eva"
- `Age`: 18, 21, 20, 19, 18
- `Major`: "Math", "Physics", "Biology", "Chemistry", "Math"
- `FirstYear`:2024, 2021, 2022, 2023, 2024

# Exercise 1: Create the data frame
```{r}
students <- data.frame(
  Name = c("Alice", "Bob", "Charlie", "David", "Eva"),
  Age = c(18, 21, 20, 19, 18),
  Major = c("Math", "Physics", "Biology", "Chemistry", "Math"),
  FirstYear = c(2024, 2021, 2022, 2023, 2024)
)
```

# Exercise 2: Count rows, columns, and dimensions
```{r}
row <- nrow(students)
col <- ncol(students)
dim(students)
```
# Exercise 3: Filter for Math majors
```{r}
students[students$Major=="Math",]

```
# Exercise 4: Filter for students older than 22
```{r}
students[students$Age>22,]
```
# Exercise 5: Filter for Math majors older than 22
```{r}
students[students$Age>22 & students$Major=="Math",]
```
# Exercise 6: Add GraduationYear column
```{r}
students$GraduationYear = c(students$FirstYear + 4)
```
# Exercise 7: Creating a Vector
```{r}
my_vec <- c(1,2,3,4,5)
my_vec_2 <- 1:5
```
# Exercise 8: Indices
```{r}
my_vec_2[1] <- 10
```
# Exercise 9: Creating functions
```{r}
my_function <- function(x,y){
  result <- x + y
  return(result)
}

my_function(2,1)
```

